Smart Drill-Down

### 2015-10-12 TableSummarization/Paper/algorithms.tex
changes
tiny change

The parameter $m_w$ stands for \textit{Max Weight}. The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$. Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, BRS is guaranteed to return $S_o$. On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by BRS does not contain $r$. BRS runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$. In practice, $\textrm{max}_{r \in S_o}W(r)$ is usually small. This is because as the size (and weight) of a rule increases, its Count falls rapidly. The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$). Thus, rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$. Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small. We later describe a strategy for setting $m_w$.

@@

\stitle{Setting parameters $W$, $k$, $m_w$:} Our system allows the user to tune the smart drill-down by adjusting a number of parameters. Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user. To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.

Parameter $k$ is the number of new rules to display upon each smart drill-down. Large values of $k$ increase the run-time quadratically, and can also overwhelm the user with too much information. On the other hand, small values of $k$ may display too little information about the table. Fortunately, the BRS algorithm is incremental in nature. That is, in order to find the best rule list of size $k+1$, it first finds the best rule-list of size $k$, and then finds another rule to add to get a rule-list of size $k+1$. Thus instead of running the algorithm with a fixed value of $k$, it can start with an empty rule-list and keep adding rules to it, displaying new rules as they are found. This search for additional rules can stop when the user issues a new smart drill-down command to the system, or manually stops the search. Alternatively, we can set a time limit (of say $5$ seconds) and display as many rules as we can find within that time limit.

$W$ is the weight function that determines which rules are interesting. This is a function specified by the user as a black box. Specifying an arbitrary function can be hard, so instead we hardcode some common Weight functions and allow the user to choose one from a drop-down menu. In addition, the user can express interest or disinterest in certain columns by telling the system of favor or ignore those columns, via a simple graphical user interface. The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.

The $m_w$ parameter lets the user trade off between the accuracy of the optimal rule-list, and running time. Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time. We cannot know the ideal value of $m_w$ in advance, but we can easily estimate it using sampling. We create a small random sample of tuples from the table, and run the BRS algorithm on it. Then the maximum weight $x$ of the output on the sample is likely to equal the maximum weight of the actual output. To account for sampling error, we can set $m_w$ to $2x$.

>>>>

The parameter $m_w$ stands for \textit{Max Weight}. The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$. Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, BRS is guaranteed to return $S_o$. On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by BRS does not contain $r$. BRS runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$. In practice, $\textrm{max}_{r \in S_o}W(r)$ is usually small. This is because as the size (and weight) of a rule increases, its Count falls rapidly. The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$). Thus, rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$. Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small. \mrj{We later describe a strategy for setting $m_w$.}

@@

\stitle{Setting parameters $W$, $k$, $m_w$:} \mrj{Our system allows the user to tune the smart drill-down by adjusting a number of parameters. Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user. To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.}

\mrj{Parameter $k$ is the number of new rules to display upon each smart drill-down. Large values of $k$ increase the run-time quadratically, and can also overwhelm the user with too much information. On the other hand, small values of $k$ may display too little information about the table. Fortunately, the BRS algorithm is incremental in nature. That is, in order to find the best rule list of size $k+1$, it first finds the best rule-list of size $k$, and then finds another rule to add to get a rule-list of size $k+1$. Thus instead of running the algorithm with a fixed value of $k$, it can start with an empty rule-list and keep adding rules to it, displaying new rules as they are found. This search for additional rules can stop when the user issues a new smart drill-down command to the system, or manually stops the search. Alternatively, we can set a time limit (of say $5$ seconds) and display as many rules as we can find within that time limit.}

\mrj{$W$ is the weight function that determines which rules are interesting. This is a function specified by the user as a black box. Specifying an arbitrary function can be hard, so instead we hardcode some common Weight functions and allow the user to choose one from a drop-down menu. In addition, the user can express interest or disinterest in certain columns by telling the system of favor or ignore those columns, via a simple graphical user interface. The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.}

\mrj{The $m_w$ parameter lets the user trade off between the accuracy of the optimal rule-list, and running time. Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time. We cannot know the ideal value of $m_w$ in advance, but we can easily estimate it using sampling. We create a small random sample of tuples from the table, and run the BRS algorithm on it. Then the maximum weight $x$ of the output on the sample is likely to equal the maximum weight of the actual output. To account for sampling error, we can set $m_w$ to $2x$. 
}

### 2015-10-12 TableSummarization/Paper/exp.tex
tiny change

\agp{Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings. The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.}

@@

\agp{
\subsubsection{Scaling properties of our algorithms}
The computation time for a smart drill-down is linear in both the table size $|T|$ and in parameter $minSS$. That is, the runtime can be written as $a \times |T| + b \times minSS$ where $a$ and $b$ are constants. In the worst-case wherer we cannot form a sample from main memory and need to re-create a sample, $a$ stands for the time taken to read data from hard disk. That is, $a \times |T|$ is the time taken to make a single pass over the table on hard disk. The constant $b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.

When the $|T|$ is small, the runtme is dominated by the $b \times minSS$ term, as seen for the Marketing Dataset in Figure~\ref{fig:minSS_speed}(a). When $|T|$ is large relative to $minSS$, like for the Census Dataset, the runtime is dominated by the $a \times |T|$ term (this is when we need to create a fresh sample from hard disk). When we have a few million tuples, we only require a few seconds. But if the dataset consisted of billions of tuples, the process of reading the table to create a sample could itself take a very long time. To counteract this, we could preprocess the dataset by down-sampling it to only a million tuples, and perform the summarization on the million tuple sample (which also effectively summarizes the billion tuple table).
}

>>>>

\mrj{Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings. The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.}

@@

\subsubsection{Scaling properties of our algorithms}
\mrj{
The computation time for a smart drill-down is linear in both the table size $|T|$ and in parameter $minSS$. That is, the runtime can be written as $a \times |T| + b \times minSS$ where $a$ and $b$ are constants. In the worst-case wherer we cannot form a sample from main memory and need to re-create a sample, $a$ stands for the time taken to read data from hard disk. That is, $a \times |T|$ is the time taken to make a single pass over the table on hard disk. The constant $b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.}

\mrj{When the $|T|$ is small, the runtme is dominated by the $b \times minSS$ term, as seen for the Marketing Dataset in Figure~\ref{fig:minSS_speed}(a). When $|T|$ is large relative to $minSS$, like for the Census Dataset, the runtime is dominated by the $a \times |T|$ term (this is when we need to create a fresh sample from hard disk). When we have a few million tuples, we only require a few seconds. But if the dataset consisted of billions of tuples, the process of reading the table to create a sample could itself take a very long time. To counteract this, we could preprocess the dataset by down-sampling it to only a million tuples, and perform the summarization on the million tuple sample (which also effectively summarizes the billion tuple table).
}

### 2015-10-12 TableSummarization/Paper/sampling.tex
tiny change

We are given a memory capacity or budget $M$, and a minimum sample size $minSS$ (both specified by the user).
The parameter $minSS$ is the minimum number of sample tuples needed
such that we can use the sample in memory 
instead of having to resort to the entire table stored
on hard disk. 
This parameter determines how accurate our count estimates will be, and also
how quickly we can return results to the user.

>>>>

We are given a memory capacity or budget $M$, and a minimum sample size $minSS$ (both specified by the user). \mrj{$M$ can simply be set to the actual available memory, while $minSS$ can easily be computed as a function of the desired estimation error.}
The parameter $minSS$ is the minimum number of sample tuples needed
such that we can use the sample in memory 
instead of having to resort to the entire table stored
on hard disk. 
This parameter determines how accurate our count estimates will be, and also
how quickly we can return results to the user. \mrj{The count estimate error is proportional to $\sqrt{\frac{1}{minSS}}$ while runtime is proportional to $minSS$.}
